Algorithms for Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2014-09-22 paper/paper.tex
Remove redundant sentence
added example to the paper

\subsection{Performance Metric}
\label{sec:perf_metric}
In order to precisely define the objective of the algrithms it is necessary to
introduce a metric that can be used to score the performance of the different
algorithms on a given problem (ensemble, budget, and deadline). The simplest
approach is to count the number of workflows in the ensemble that each algorithm
is able to complete within the budget before the deadline, but this metric does
not account for the priority-based utility function specified by the user. Using
the counting approach, a less efficient algorithm may be able to complete a
large number of low-priority workflows by executing the smallest workflows
first. In order to account for the priority, we use an exponential
scoring defined as:
$$ Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}} $$
where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that
the highest-priority workflow has $Priority(w)=0$, the next highest workflow has
$Priority(w)=1$, and so on. This exponential scoring function gives the highest
priority workflow a score that is higher than all the lower- priority workflows
combined:
$$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
This scoring is consistent with our definition of the problem, which is to
complete as many workflows as possible, according to their priorities, given a
set budget and deadline.

@@

DPDS guarantees that tasks from lower priority workflows are always deferred
when higher-priority tasks are available, but lower-priority tasks can still
occupy idle VMs when higher-priority tasks are not available. Since there is
no preemption, long-running low-priority tasks may delay the execution of
higher-priority tasks. In addition, tasks from low priority workflows may be
executed even though there is no chance that those workflows will be completed
within the current budget and deadline. Figure~\ref{fig:algorithm-example-dpds}
shows an example schedule generated using the DPDS algorithm. The figure
illustrates how tasks from lower priority workflows backfill idle VMs when
tasks from higher priority workflows are not ready for execution.

@@

The previous dynamic (online) algorithms make provisioning and scheduling
decisions at runtime. By contrast, the SPSS algorithm creates a provisioning
and scheduling plan before running any workflow tasks. This enables SPSS to
start only those workflows that it knows can be completed given the deadline and
budget constraints, and eliminates any waste that may be allowed by the dynamic
algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
changes in the environment (see Sections~\ref{sec:variances} and
\ref{sec:delays}).

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}

@@

An example schedule generated by SPSS (Figure~\ref{fig:algorithm-example-spss})
shows how SPSS tends to start many workflows in parallel, running each workflow
over a longer period of time on only a few VMs to minimize cost. In comparison,
the dynamic algorithms tend to run one workflow at a time across many VMs in
parallel.

>>>>

\subsection{Performance Metric}
\label{sec:perf_metric}
In order to precisely define the objective of the algrithms it is necessary to
introduce a metric that can be used to score the performance of the different
algorithms on a given problem (ensemble, budget, and deadline). The simplest
approach is to count the number of workflows in the ensemble that each algorithm
is able to complete within the budget before the deadline, but this metric does
not account for the priority-based utility function specified by the user. Using
the counting approach, a less efficient algorithm may be able to complete a
large number of low-priority workflows by executing the smallest workflows
first. In order to account for the priority, we use an exponential
scoring defined as:
\begin{equation}
\label{eq:score}
Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
\end{equation}
where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that
the highest-priority workflow has $Priority(w)=0$, the next highest workflow has
$Priority(w)=1$, and so on. This exponential scoring function gives the highest
priority workflow a score that is higher than all the lower- priority workflows
combined:
$$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
This scoring is consistent with our definition of the problem, which is to
complete as many workflows as possible, according to their priorities, given a
set budget and deadline.

@@

DPDS guarantees that tasks from lower priority workflows are always deferred
when higher-priority tasks are available, but lower-priority tasks can still
occupy idle VMs when higher-priority tasks are not available. Since there is
no preemption, long-running low-priority tasks may delay the execution of
higher-priority tasks. In addition, tasks from low priority workflows may be
executed even though there is no chance that those workflows will be completed
within the current budget and deadline.

@@

The previous dynamic (online) algorithms make provisioning and scheduling
decisions at runtime. By contrast, the SPSS algorithm creates a provisioning
and scheduling plan before running any workflow tasks. This enables SPSS to
start only those workflows that it knows can be completed given the deadline and
budget constraints, and eliminates any waste that may be allowed by the dynamic
algorithms.

@@

\subsection{Illustrative example}

To illustrate how the algorithms perform, we parepared an artificial ensemble 
of three workflows, as seen in Fig.~\ref{fig:example-ensemble}. 
The deadline is set to 6 hours and the budget to \$18. In this example we assume
that there are no VM provisioning delays, the runtime estimates are accurate, 
there are no failures and the cost of 1 VM-hour is \$1.
For dynamic algorithms we assume that the low utilization threshold ($u_l$) for deprovisioning is 50\%
and that the autoscaling parameter $v_{max} = 1$.

\begin{figure}[htb] 
  \centering
  \includegraphics[width=0.80\columnwidth]{figures/example-ensemble}
  \caption{Artificial ensemble consisting of three workflows \{a, b, c\}, with priorities \{1, 2, 3\}.
  Task labels indicate estimated execution time in minutes, 
  so e.g. task $a.70$ has estimated execution time of 70 minutes.}
  \label{fig:example-ensemble}
\end{figure}

The resulting schedule produced by DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-dpds}.
The algorithm at the beginning estimates that it needs 3 VM instances by dividing the budget by the deadline ($18 / 6 = 3$). 
Next, it schedules all the ready tasks in the priority order, so $a.70$ is started first, 
and immediately $b.70$ and $c.60$ are started, since no other tasks of workflow a are ready. 
When task $a.60$ completes, there are again other no higher priority tasks ready, 
so the next task from workflow $c$ can be started ($c.45$). Since there is no preemption,
this results in workflow $a$ being delayed by the lower priority workflow $c$.
Finally, the outcome of this dynamic scheduling policy is that none of the workflows completes within the deadline.
Actually, the tasks $b.100$, $c.65$ and $a.160$ are not finished since all the VMs 
are terminated due to running out of budget. 
The exponential score achieved by the algorithm computed according to equation~\ref{eq:score} is 0.
On the other hand, we can observe that the utilization of the VMs is kept on the level of 100\% throughout the 
whole execution time, which may be considered an advantage in some cases.

\begin{figure}[htb] 
  \centering
  \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-dpds}
  \caption{Schedule produced by DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
  White boxes indicate VM instances provisioned for their hourly time slots, 
  and solid boxes indicate running tasks.
  None of the workflows completes within the deadline.}
  \label{fig:algorithms-example-dpds}
\end{figure}

The resulting schedule produced by WA-DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-wadpds}.
The schedule is generated in a similar way as in DPDS algorithm, with the one notable difference. 
After scheduling tasks $a.70$ and $b.70$, the algorithm estimates whether it has enough budget to 
admit workflow $c$. The answer is that based on the sum of task runtimes there is no room for workflow $c$,
so the whole workflow is rejected. As a consequence, task $a.100$ can start at time $t=70$ minutes
thus workflow $a$ is not delayed as in the case of DPDS. One can also observe that $VM.2$ and $VM.3$ remain idle for
one hour, which is the consequence of the conservative provisioning settings we assume in this example.  
This means that VM utilization at the level of 2/3 does not trigger deprovisioning in this case, and the maximum autoscaling
factor set to 1 prevents from provisioning of new VMs over the initial number even if utilization is 100\%.
In the resulting schedule only workflow $a$ completes, so the score is $2^{-1} = 0.5$, which is better than DPDS.

\begin{figure}[htb] 
  \centering
  \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-wadpds}
  \caption{Schedule produced by WA-DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
  Workflow a completes within the deadline.}
  \label{fig:algorithms-example-wadpds}
\end{figure}

The resulting schedule produced by SPSS algorithm is shown in Fig.~\ref{fig:algorithms-example-spss}.

\begin{figure}[htb] 
  \centering
  \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-spss}
  \caption{Schedule produced by DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
  Workflows a and c complete within the deadline.}
  \label{fig:algorithms-example-spss}
\end{figure}

The example given above illustrates the main concepts of the algorithms using the artificial ensemble.
Two additional example schedules generated using our simulator are shown in 
Figure~\ref{fig:examples-simulated}. 
The schedule generated using DPDS (Figure~\ref{fig:algorithm-example-dpds})
illustrates how tasks from lower priority workflows backfill idle VMs when
tasks from higher priority workflows are not ready for execution.
On the other hand, the schedule generated by SPSS (Figure~\ref{fig:algorithm-example-spss})
shows how SPSS tends to start many workflows in parallel, running each workflow
over a longer period of time on only a few VMs to minimize cost. In comparison,
the dynamic algorithms tend to run one workflow at a time across many VMs in
parallel.

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}
  \label{fig:examples-simulated}

@@
